---
title: "US COVID"
date: "04/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### INTRODUCTION
Hey welcome to my Covid 19 report and analysis. COVID-19 is a dangerous virus that has killed millions of  people all over the world. This is a very serious pandemic and it is still going on right now. This report will be focusing on Covid-19 in the United States. The data that I will be using for my report can be found on github, the link is provided here : "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series". This is a time series data set that has many attributes. This is a csv dataset that is updated daily regarding the number of cases, the number of deaths and the number of recovery. This data is from Johns Hopkins University. Thanks to them we are able to do our own reports and analysis. 

What I will be focusing on in my reports is the correlation between the number of cases and the number of deaths regarding Covid-19 in the United States. I will also be looking in the state of Califorina to see how the cases are matched up against compared to the cases in New York. 

### DATA PREPERATION 
 
 
```{r}
library(tidyverse)
library(lubridate)
confirmed_link <- c("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")
confirmed_df <- read_csv(confirmed_link)
tail(confirmed_df)
us_death <- c("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
us_deathdf <- read_csv(us_death)
us_cases <- confirmed_df %>% pivot_longer(cols = -(UID:Combined_Key),names_to = "date",values_to = "cases") %>% select(Admin2:cases) %>% mutate(date = mdy(date)) %>% select(-c(Lat,Long_))
us_deaths <- us_deathdf %>% pivot_longer(cols = -(UID:Population),names_to = "date",values_to = "deaths") %>% select(Admin2:deaths) %>% mutate(date = mdy(date)) %>% select(-c(Lat,Long_))
us_df<-us_cases %>% full_join(us_deaths)
us_by_state <- us_df %>% group_by(Province_State,Country_Region,date) %>% summarize(cases = sum(cases),deaths = sum(deaths),Population=sum(Population)) %>% mutate(deaths_per_million = deaths * 1000000 / Population) %>% select(Province_State,Country_Region,date,cases,deaths,deaths_per_million,Population) %>% ungroup()
us_totals <- us_by_state %>% group_by(Country_Region,date) %>% summarize(cases = sum(cases), deaths=sum(deaths),Population = sum(Population)) %>% mutate(deaths_per_million = deaths * 1000000 / Population) %>% select(Country_Region, date, cases,deaths,deaths_per_million, Population) %>% ungroup()
us_state_totals <- us_by_state %>% group_by(Province_State)%>% summarize(deaths=max(deaths),cases=max(cases), population= max(Population),cases_per_thou=1000*cases/population,deaths_per_thou=1000* deaths/population) %>% filter(cases>0,population>0)

```



### Graph Preperation 

```{r}
p1 <- ggplot(us_totals, aes(date)) +     
geom_line(aes(y=cases), colour="red") +
ggtitle("Covid-19 in US")
state <- "California"
cali <- us_by_state %>% filter(Province_State== state) %>% ggplot(aes(x=date,y=cases))+geom_line(aes(color="cases"))+geom_point(aes(color="cases"))+ geom_line(aes(y=deaths,color="deaths"))+geom_point(aes(y=deaths,color="deaths"))+scale_y_log10()+ labs(title="COVID-19 in California")
state2 <- "New York"
ny <-  us_by_state %>% filter(Province_State== state2) %>% ggplot(aes(x=date,y=cases))+geom_line(aes(color="cases"))+geom_point(aes(color="cases"))+ geom_line(aes(y=deaths,color="deaths"))+geom_point(aes(y=deaths,color="deaths"))+scale_y_log10()+ labs(title="COVID-19 in New York")
mod <- lm(deaths_per_thou ~ cases_per_thou,data=us_state_totals)
pred_ <- us_state_totals %>% mutate(pred=predict(mod))
model_graph <- pred_ %>% ggplot() +geom_point(aes(x=cases_per_thou,y=deaths_per_thou),color="green")+ geom_point(aes(x=cases_per_thou,y = pred),color="red")
```


### GRAPHS

Here are some basic visualization from the state of California.
```{r}
plot(cali)
plot(ny)
```

### MODELS 
Here is a Linear regression model of cases per thousand versus deaths per thousands from all the states in United States.

```{r}
plot(model_graph)
```


### CONCLUSION AND BIAS 

So from the visualization you can tell when cases goes up the death also goes up so there is a correlation between them both. Some bias that are applicable in this reports is that not all cases and deaths willbe reported. Not all cases are covid related but maybe a different type of virus and all the deaths are not covid deaths. Since covid is a new virus and there are not information or reaserach done on this virus its hard to tell if the cases and deaths were Coivd-19 related. 
